The Effect of Events on Discussion in the Australian Federal Parliament (1901--2017)

### 2018-10-22 outputs/docs/events/paper.Rmd
MA edits

In this paper we examine text records of what was said in the Australian Federal Parliament. We create a dataset based on all sitting days from 1901 through to 2017. Although we drop a handful of days for various reasons, we have records for 7,869 days in the House of Representatives and 6,682 days in the Senate. Our dataset consists of counts of every word that occurs at least thrice, after we convert all words to lower-case, remove numbers and punctuation, join commonly co-located words, and remove stop words.

To reduce the dimensionality of our dataset we use a topic model that allows correlation between topics. We then analyse the topics using a Bayesian hierarchical Dirichlet model to examine changes at various types of events. These events include: changes in government; elections; and other significant events such as an economic recession or terrorism.

@@

The recent digitisation of the Canadian Hansard, @BeelenEtc2017, allowed @RheaultCochran2018 to examine ideology and party polarisation in Britain and Canada. In the UK, @Duthie2016 analysed Hansard records to examine which politicians made supportive or aggressive statements toward other politicians between 1979 and 1990 and @PetersonSpirling2018 examined polarisation. And as digitisation methods improve older UK records can be analysed, for instance @Dimitruk2018 considers the effect of estate bills on prorogations in seventeenth century England. In New Zealand, @Curran2017 modelled the topics discussed between 2003 and 2016, and @Graham2016 examined unparliamentary language between 1890 and 1950. And in the US @GentzkowShapiroTaddy2018 examined congressional speech records from 1873 to 2016 to find that partisanship has risen in the past few decades.

@@

$$\theta_d \sim \mbox{Logistic Normal}(\mu, \Sigma)$$
This is, the main difference between the CTM and LDA is that it replaces the assumption of the Dirichlet distribution with a more flexible logistic Normal distribution. That distribution can incorporate a covariance structure across the topics. The remainder of the steps of the document generating process are identical to LDA.

However, the replacement of the Dirichlet distribution, which is conjugate to the multinomial, with the logistic Normal distribution adds a level of computational complexity to CTM. The posterior distributions of the parameters of interest ($\beta_{1:K}, \theta_{1:D}, z_{1:D, 1:N}$) can no longer be obtained using standard simulation techniques such as Gibbs Sampling. @BleiLafferty2007 develop a fast variational inference procedure for estimating the CTM. CTM itself has also been extended by @RobertsStewartAiroldi2016 as part of their work on Structural Topic Models. The main difference is to add a covariate to $\mu$ which allows consideration of additional information.

@@

The goal of the model is to relate these proportions to government $g$ at time $d$, and also the days since the most recent election, $e$, while account for underlying time trends. The mean parameters $\mu_{sp}$ are modelled on the log scale as

@@

$$
\alpha_{e[s],d,p} = \rho_{e[s],p} \cdot \alpha_{e[s],d-1,p}
$$
The value of the initial effect, $\alpha_{e[s],0,p}$, and the AR(1) term, $\rho$, both have non-informative priors:

$$
\alpha_{e[s],0,p} \sim \mbox{Normal}(0, 100)
$$
$$
\rho_{e[s],p} \sim \mbox{Uniform}(0,1)
$$
We model the underlying time trend in topics using splines regression. The intuition behind this term is to capture the underlying non-linear trend in topic distributions over time, which is caused by large-scale structural changes in the economy, and Australian society and culture. The $x_{s,k}$ for $k = 1,2,\dots, K$ are the value of cubic basis splines for sitting period $s$ at knot point $k$. We place knot points every five sitting periods as this is the average length of time for a government to sit (**CHECK**). Non-informative priors are placed on the splines coefficients:
$$
\beta_{p,k} \sim \mbox{Normal}(0, 100)
$$

@@

We run the model in JAGS using the \texttt{rjags} package of @Plummer2018. __[An illustration of the validity of our analysis model using simulated data is in Appendix \@ref(modeldetailsandsimulation).] __

@@

# Results
We are interested in considering the effect of various political and other events on what is talked about in the Australian Federal Parliament. Political events are those related to a change of government or an election. As Australia has a parliamentary system, it is possible for the government to change without an election. We define a government based on who is the prime minister, and do not distinguish between terms or cabinet composition as is sometimes done. If a person was prime minister more than once then these are considered separately. If the mean parameter for a government or election period is outside the 95 per cent credible interval, then we consider that period to be different.

We summarise our results in terms of elections Table \@ref(tab:electionsresults) and in terms of governments in Table \@ref(tab:governmentsresults). These tables focus on elections and governments that were different to the ones that preceded them.

@@

In Figures \@ref(fig:mugov) and \@ref(fig:muelection) we focus on certain topics to illustrate differences between governments and elections, respectively.

```{r mugov, cache = TRUE, echo = FALSE, fig.cap = "Government level effects on selected topics", out.width = '100%'}
knitr::include_graphics("../../figures/mu_gov.pdf")
```

@@

The topics that we have focused on here are:
12, 17, 22, 23 which have to do with conflict, defence and security; 13 and 32 which have to do with commerce and trade; 24 and 33 which have to do with the budget, tax and spending; 29 which is to do with education and 39 which has to do with health.

@@

We summarise the days where parliamentary discussion was significantly different from the rest prevailing in that week in Table \@ref(tab:keyeventsresultstable).

@@

Finally, we summarise our results over time by looking at the standard deviation of sitting period estimates, around the mean election effect (Figure \@ref(fig:sigmaelection))

@@

# Discussion
In this paper we consider what was said in the Australian Federal Parliament between 1901 and 2017. We download and parse PDFs of Hansard to create a new dataset of text. We use a correlated topic model to group the parliamentary discussion into topics to reduce the dimensionality, and then analyse the effect of various events on the distribution of these topics using a Bayesian hierarchical Dirichlet model. In general we find that changes in government change the distribution of topics discussed in parliament, but that most elections did not. We find that significant events such as 9/11 and the Bali Bombings had a substantial effect, but that with certain exceptions, economic events did not.

We identify differences between neighbouring governments and between neighbouring elections based on calculating 95 per cent credible intervals based on posterior samples. When these do not overlap we consider that the model finds a significant difference between either the neighbouring governments or elections. One advantage of our model over using the STM is that we can restrict the effect of an election to be monotonically decreasing. This allows us to identify differences between government and election effects even when there is a one-term government.

Of the 36 different governments over this period, we find that 14 of them are significantly different to the government that preceded them. However, two of these results -- the significance of the Page  Government and the first Menzies Government -- are likely due to the short length of the Page Government and should be ignored, leaving only 12. The earliest of these are the Second and Third Fisher Governments, which were different to the Third Deakin Government and the Cook Government respectively. To a certain extent this is likely tied up with changes due to World War I.

@@

One advantage of our analysis model, compared with using the STM approach is that we can create a measure that is equivalent to testing for outliers in a model where the underlying variables were not latent. We estimate sitting-period level-effects (essentially a median for each topic by sitting period). This can be thought of as as equivalent to a 'actual' level, and then the points for we estimate are the 'estimated' points. The difference defines a measure that can be thought of as essentially a residual and we then identify days where the is at least a 99 per cent difference. This approach means that the model generates dates that are interesting without us having to specify interesting dates.

@@

@@

**[Add discussions about weakness of the model.]** The main weakness of this approach is that it does not appropriately propagate uncertainty and future research should improve on this.

Additionally, the way that we identify unusual periods could be improved. e.g. What happens if an event happens while parliament not sitting then it's not clear how whether that is going to be identified. And if an event happens in the middle of an sitting period, but it's really important and changes the discussion of the rest of the period then again that might not be different because of the average comparison.

>>>>

In this paper we examine text records of what was said in the Australian Federal Parliament. We create a dataset based on all sitting days from 1901 through to 2017. We have records for 7,869 days in the House of Representatives and 6,682 days in the Senate. Our dataset consists of counts of every word that occurs at least thrice, after we convert all words to lower-case, remove numbers and punctuation, join commonly co-located words, and remove stop words.

We systematically analyse the text using two main statistical techniques. To reduce the dimensionality of our dataset we use a topic model that allows correlation between topics. We then analyse the topics using a Bayesian hierarchical Dirichlet model to examine changes at various types of events. These events include: changes in government; elections; and other significant events such as an economic recession or terrorism.

@@

The recent digitisation of the Canadian Hansard, @BeelenEtc2017, allowed @RheaultCochran2018 to examine ideology and party polarisation in Britain and Canada. In the UK, @Duthie2016 analysed Hansard records to examine which politicians made supportive or aggressive statements toward other politicians between 1979 and 1990 and @PetersonSpirling2018 examined polarisation. And as digitisation methods improve older UK records can be analysed, for instance @Dimitruk2018 considers the effect of estate bills on prorogations in seventeenth century England. In New Zealand, @Curran2017 modelled the topics discussed between 2003 and 2016, and @Graham2016 examined unparliamentary language between 1890 and 1950. And in the US, @GentzkowShapiroTaddy2018 examined congressional speech records from 1873 to 2016 to find that partisanship has risen in the past few decades.

@@

$$\theta_d \sim \mbox{Logistic Normal}(\mu, \Sigma)$$
That is, the main difference of CTM over LDA is that it replaces the assumption of the Dirichlet distribution with a more flexible logistic multivariate Normal distribution. This distribution can incorporate a covariance structure across the topics. The remainder of the steps of the document generating process are identical to LDA.

However, the replacement of the Dirichlet distribution with the logistic multivariate Normal distribution adds a level of computational complexity to CTM. The posterior distributions of the parameters of interest ($\beta_{1:K}, \theta_{1:D}, z_{1:D, 1:N}$) can no longer be obtained using standard simulation techniques such as Gibbs Sampling. @BleiLafferty2007 develop a fast variational inference procedure for estimating the CTM. CTM itself has also been extended by @RobertsStewartAiroldi2016 as part of their work on Structural Topic Models. The main difference is to add a covariate to $\mu$ which allows consideration of additional information.

@@

The goal of the model is to relate these proportions to government $g$ at time $d$, and also the days since the most recent election, $e$, while account for underlying time trends. The mean parameters $\mu_{s,p}$ are modelled on the log scale as

@@

$$
\alpha_{e[s],d,p} = \rho_{e[s],p} \cdot \alpha_{e[s],d-1,p}
$$
One advantage of our model over using the STM is that we can restrict the effect of an election to be monotonically decreasing. This allows us to identify differences between government and election effects even when there is a one-term government. The value of the initial effect, $\alpha_{e[s],0,p}$, and the AR(1) term, $\rho$, both have non-informative priors:

$$
\alpha_{e[s],0,p} \sim \mbox{Normal}(0, 100)
$$
$$
\rho_{e[s],p} \sim \mbox{Uniform}(0,1)
$$
We model the underlying time trend in topics using splines regression. The intuition behind this term is to capture the underlying non-linear trend in topic distributions over time, which is caused by large-scale structural changes in the economy, and Australian society and culture. The $x_{s,k}$ for $k = 1,2,\dots, K$ are the value of cubic basis splines for sitting period $s$ at knot point $k$. We place knot points every five sitting periods as this is the average length of time for a government to sit. Non-informative priors are placed on the splines coefficients:
$$
\beta_{p,k} \sim \mbox{Normal}(0, 100)
$$

@@

We run the model in JAGS using the \texttt{rjags} package of @Plummer2018.

<!-- __[An illustration of the validity of our analysis model using simulated data is in Appendix \@ref(modeldetailsandsimulation).] __ -->

@@

# Results
We are interested in considering the effect of various political and other events on what is talked about in the Australian Federal Parliament. Political events are those related to a change of government or an election. As Australia has a parliamentary system, it is possible for the government to change without an election. We define a government based on who is the prime minister, and do not distinguish between terms or cabinet composition as is sometimes done. If a person was prime minister more than once then these are considered separately.

We identify differences between neighbouring governments and between neighbouring elections based on calculating 95 per cent credible intervals based on posterior samples. When these do not overlap we consider that the model finds a significant difference between either the neighbouring governments or elections.

We summarise our results in terms of governments in Table \@ref(tab:governmentsresults) and in terms of elections in Table \@ref(tab:electionsresults). These tables focus on elections and governments that were different to the ones that preceded them.

@@

In Figures \@ref(fig:mugov) and \@ref(fig:muelection) we focus on certain topics to illustrate differences between governments and elections, respectively. In the graphs, the points show the estimated value of $\alpha_g$ and $\alpha_e$, respectively, for each of the topics specified. The error bars represent 95 per cent Bayesian credible intervals. The topics that we have focused on here are:
12, 17, 22, 23 which have to do with conflict, defence and security; 13 and 32 which have to do with commerce and trade; 24 and 33 which have to do with the budget, tax and spending; 29 which is to do with education and 39 which has to do with health.

```{r mugov, cache = TRUE, echo = FALSE, fig.cap = "Government level effects on selected topics.", out.width = '100%'}
knitr::include_graphics("../../figures/mu_gov.pdf")
```

@@

@@

We estimate sitting-period level-effects (essentially a mean for each topic by sitting period). The difference between this mean distribution and a particular day's topic distribution defines a measure that can be thought of as essentially a residual which allows us to identify outlying days. This approach means that the model generates dates that are interesting without us having to specify interesting dates.

More specifically, we define a day to be 'outlying' or 'different-to-expected' if the topic distribution on that particular day is more than three standard deviations different to the mean topic distribution for the relevant sitting period. Table \@ref(tab:keyeventsresultstable) summarises the days where parliamentary discussion was significantly different from the rest prevailing in that week.

@@

Finally, Figure \@ref(fig:sigmaelection) illustrates the estimated standard deviation of topic distributions, around the mean election effect, over time. In general, the estimates of standard deviation are increasing over time, suggesting that discussed topics are more variable in more recent election periods compared to historical periods.

@@

# Discussion
Of the 36 different governments over this period, we find that 14 of them are significantly different to the government that preceded them. However, two of these results -- the significance of the Page  Government and the first Menzies Government -- are likely due to the short length of the Page Government and should be ignored, leaving only 12. The earliest of these are the Second and Third Fisher Governments, which were different to the Third Deakin Government and the Cook Government respectively. To a certain extent this is likely tied up with changes due to World War I.

@@

One advantage of our analysis model, compared with using the STM approach is that we can create a measure that is equivalent to testing for outliers in a model where the underlying variables were not latent.

@@

# Conclusion
In this paper we consider what was said in the Australian Federal Parliament between 1901 and 2017. We download and parse PDFs of Hansard to create a new dataset of text. We use a correlated topic model to group the parliamentary discussion into topics to reduce the dimensionality, and then analyse the effect of various events on the distribution of these topics using a Bayesian hierarchical Dirichlet model. In general we find that changes in government change the distribution of topics discussed in parliament, but that most elections did not. We find that significant events such as 9/11 and the Bali Bombings had a substantial effect, but that with certain exceptions, economic events did not.

@@

In terms of the analysis model which relates the topic distributions with events, there are several limitations to the model. Firstly, we are assuming that the effect of a particular government is constant across the whole period. In addition, we assume that the effect of elections is monotonically decreasing across days since election. In future work we will investigate different functional forms on both of these effects, and in particular try to allow for elections to have a 'lead-up' effect.

The way that we identify unusual periods could also be improved. We defined sitting periods in a constant fashion across the whole dataset time frame, but the nature of how long parliaments sit for has most likely changed. In addition, more work needs to be done on how to identify outlying events. For example, it is not clear if an important event that occurs outside a sitting period would be identified. And if an event happens in the middle of an sitting period, it may have a large effect on the overall mean, such specific days are not identified as significantly outlying.

Finally, the current modeling and analysis set-up is a two stage process: we take the output of a topic model, and use this as the input to a second model. However, this approach does not appropriately propagate uncertainty in topic distribution estimation. Future methodological work will focus on how to combine these two modeling steps by extending the STM approach into a more flexible framework.
